---
title: "Tuning RF with 'mtry'"
author: "Matthew Curcio"
date: "March 27, 2019"
output: html_document
---

## Title: Tuning_RF_with_mtry.rmd

#### Summary: Parameter Tuning (mtry) for Random Forest when Using caret Package

#### Library 1/2:
```{r library1, message=FALSE}
cache = TRUE
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "doMC",  "caret", "randomForest")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```

#### Import Data:
```{r import}
cache = TRUE
setwd("~/Dropbox/Oxy-RF/4_RF_Tuning_w_mtry")
test_harness_paa <- read_csv("test_harness_paa.csv",
                             col_types = cols(TotalAA = col_skip(),
                                              id = col_skip()))
```

#### Convert 'Class' To Factor:
Convert Class(numerical) to Factor of 7 Protein Classes(Prot_Class)
```{r factor}
cache = TRUE
class_factor <- as.factor(test_harness_paa$Class)

typeof(test_harness_paa)
class(class_factor)
```

#### Partition Testing and Training datasets:
```{r partition}
cache = TRUE
set.seed(1000)

index <- createDataPartition(test_harness_paa$Class, p = 0.8, list = FALSE)
training_set <- test_harness_paa[ index,]
testing_set  <- test_harness_paa[-index,]

preProcValues <- preProcess(training_set, method = c("center", "scale"))

train_transformed <- predict(preProcValues, training_set)
test_transformed  <- predict(preProcValues, testing_set)
```

#### Train control for 10 X Cross-validation with 5 reps:
```{r control1}
cache = TRUE
set.seed(1000)
train_control <- trainControl(method = "repeatedcv", 
                              number = 10,
                              repeats = 5,
                              verboseIter = TRUE,
                              allowParallel = TRUE,
                              summaryFunction = multiClassSummary)
```

#### 'mtry' grid tuning:
```{r train2}
cache = TRUE
set.seed(1000)
registerDoMC(cores = 3)

start_time <- Sys.time() # Start timer

my_grid1 <- expand.grid(mtry = 1:4)

rf1 <- train(Class ~ .,
             data = train_transformed,
             method = "rf",
             metric = "Accuracy",
             tuneGrid = my_grid1,
             trControl = train_control)

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```

#### View Results:
```{r view}
rf1
```

#### Library 2/2:
```{r library2, message=FALSE, warning=FALSE, include=FALSE, message=FALSE}
library(magrittr)
library(dplyr)
library(tidyverse)
```

#### Plotting:
```{r plot}
cache = TRUE

my_plot <- function(model) {
    theme_set(theme_minimal())
    u <- model$results %>%
        select(mtry, Accuracy, Kappa, Mean_F1, Mean_Sensitivity,
               Mean_Specificity,Mean_Pos_Pred_Value, Mean_Neg_Pred_Value, 
               Mean_Precision, Mean_Recall, Mean_Detection_Rate) %>%
        gather(a, b, -mtry)
    
    u %>% ggplot(aes(mtry, b)) + geom_line() + geom_point() + 
        facet_wrap(~ a, scales = "free") + 
        labs(x = "Number of mtry", y = NULL, 
             title = "The Relationship between Model Performance and mtry")
}

rf1 %>% my_plot()
```


Machine Settings:
```{r}
Sys.info()[c(1:3,5)]
```
```{r}
sessionInfo()
```
EOF
