---
title: "Tuning_RF_Using_ntree.rmd"
author: "Matthew Curcio"
email: "mockrun@gmail.com"
date: "March 27, 2019"
output: 
  html_document: 
    toc: yes
    toc_float: true
    collapsed: FALSE
---

## Title: Tuning RF Using ntree

#### Summary: Parameter Tuning (ntree) for Random Forest when Using caret Package

#### Library:
```{r library, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

Libraries = c("readr", "doMC",  "caret", "randomForest", "ggplot2")

# Install if not present
for(p in Libraries){
    if (!require(p, character.only = TRUE)) { install.packages(p) }
    library(p, character.only = TRUE)
}
```

#### Import Data:
```{r import}
setwd("~/Dropbox/Oxy-RF/5_RF_Tuning_w_ntree")
test_harness_paa <- read_csv("test_harness_paa.csv",
                             col_types = cols(TotalAA = col_skip(),
                                              id = col_skip()))
```

#### Convert 'Class' To Factor:
Convert Class(numerical) to Factor of 7 Protein Classes(Prot_Class)
```{r factor}
Class <- as.factor(test_harness_paa$Class)

typeof(test_harness_paa)
class(Class)
```

#### Partition Testing and Training datasets:
```{r partition}
set.seed(1000)

index <- createDataPartition(test_harness_paa$Class, p = 0.8, list = FALSE)
training_set <- test_harness_paa[ index,]
testing_set  <- test_harness_paa[-index,]

preProcValues <- preProcess(training_set, method = c("center", "scale"))

training_set <- predict(preProcValues, training_set)
```


#### A. RF 'mtry=1,2,3,4' tuning:
```{r train}
set.seed(1000)
registerDoMC(cores = 3)

start_time <- Sys.time() # Start timer

# How many columns to select in each bootstrap sample?
t_grid <- expand.grid(mtry= c(1, 2, 3, 4))

model.rf <- train(Class ~ .,
                  data = training_set,
                  method = "rf",
                  ntree = 100, # How many trees to grow in total?
                  tuneGrid = t_grid)

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```

#### B. RF 'mtry=1,2,3,4' Tuning Results:
```{r}
print(model.rf)
```

#### C. RF  ntree = 500 tuning:
```{r}
set.seed(1000)
registerDoMC(cores = 3)

start_time <- Sys.time() # Start timer

model.rf <- train(Class ~ .,
                  data = training_set,
                  method = "rf",
                  ntree = 500)

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```

#### D. Run Correlation Analysis to investigate for multicollinearity:
```{r}
# 
trainingCor <- findCorrelation(cor(training_set[,-1]), cutoff = 0.60)

trainingCor
```


```{r}
rf <- randomForest(Class ~ ., data = training_set, ntree = 100, mtry = 2)
rf
```



Machine Settings:
```{r}
Sys.info()[c(1:3,5)]
```
```{r}
sessionInfo()
```
EOF
