---
title: "Intial Scouting-Random Forest of AA USING Tesst-Harness.rmd"
author: "Matthew Curcio"
date: "March 20, 2019"
output: html_document
---

## Title: Intial Scouting-Random Forest of AA USING Tesst-Harness

### Summary: Using the 'caret' package from Max Kuhn and the Test-Harness file I will run random forest to classify the 7 protein groups.

#### Libraries:
```{r libraries, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "doMC", "caret")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```


#### Import Data:
```{r}
test_harness <- read_csv("test_harness_paa.csv")
```


#### Convert 'Class' To Factor
Convert Class(numerical) to Factor of 7 Protein Classes(Prot_Class)
```{r factor}
prot_class <- as.factor(complete_aa$Class)

typeof(prot_class)
class(prot_class)
```


#### Partition Testing and Training datasets
```{r}
set.seed(1000)

index <- createDataPartition(complete_aa$Class, p = 0.8, list = FALSE)
training_set <- complete_aa[ index,]
testing_set  <- complete_aa[-index,]
```


#### RF model
```{r}
start_time <- Sys.time() # Start timer

registerDoMC(cores=3)
modFit <- train(training_set$Class ~ ., data = training_set, method = "rf")

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```


Machine Settings:
```{r}
Sys.info()[c(1:3,5)]
```
```{r}
sessionInfo()
```
EOF