---
title: "Intial Scouting-Random Forest of AA USING Test-Harness.rmd"
author: "Matthew Curcio"
date: "March 20, 2019"
output: html_document
---

## Title: Intial Scouting-Random Forest of AA Using Test-Harness

### Summary: Using the 'caret' package from Max Kuhn and the Test-Harness file I will run random forest to classify the 7 protein groups.

#### Libraries:
```{r libraries, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "doMC", "caret")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```


#### Import Data:
```{r import}
test_harness_paa <- read_csv("test_harness_paa.csv")
```


#### Convert 'Class' To Factor:
Convert Class(numerical) to Factor of 7 Protein Classes(Prot_Class)
```{r factor}
class_factor <- as.factor(test_harness_paa$Class)

typeof(test_harness_paa)
class(class_factor)
```


#### Pre-processing: 
```{r preproc}
test_harness_paa <- test_harness_paa[,-c(2:3)] # More later
```


#### Partition Testing and Training datasets:
```{r partition}
set.seed(1000)

index <- createDataPartition(test_harness_paa$Class, p = 0.8, list = FALSE)
training_set <- test_harness_paa[ index,]
testing_set  <- test_harness_paa[-index,]
```


#### RF model training
```{r train}
start_time <- Sys.time() # Start timer

registerDoMC(cores=3)
modFit <- train(Class ~ ., 
                data = training_set,
                method = "rf")

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```


#### Model fitting output:
```{r}
modFit
```

#### Testing 'mtry' with 1:5
```{r}

```


Machine Settings:
```{r settings}
Sys.info()[c(1:3,5)]
```
```{r session}
sessionInfo()
```
EOF