---
title: "Random-Forest-of-AA-on-Test-Harness_w_Scale_Center.rmd"
author: "Matthew Curcio"
date: "March 20, 2019"
output: html_document
---

## Title: Random-Forest-of-AA-on-Test-Harness_w_Scale_Center

### Summary: Using the 'caret' package from Max Kuhn and the Test-Harness file I will run random forest to classify the 7 protein groups. 

This script has **preProcess = c("center","scale")**.

#### Libraries:
```{r libraries, message=FALSE}
cache = TRUE
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "doMC", "caret")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```

#### Import Data:
```{r import}
test_harness_paa <- read_csv("test_harness_paa.csv")
```

#### Convert 'Class' To Factor:
Convert Class(numerical) to Factor of 7 Protein Classes(Prot_Class)
```{r factor}
class_factor <- as.factor(test_harness_paa$Class)

typeof(test_harness_paa)
class(class_factor)
```

#### Data Cleaning: 
```{r clean}
test_harness_paa <- test_harness_paa[,-c(2:3)]
```

#### Partition Testing and Training datasets:
```{r partition}
cache = TRUE
set.seed(1000)

index <- createDataPartition(test_harness_paa$Class, p = 0.8, list = FALSE)
training_set <- test_harness_paa[ index,]
testing_set  <- test_harness_paa[-index,]
```

#### Implement Pre-Processing to compare training
```{r train}
cache = TRUE
registerDoMC(cores=3)

start_time <- Sys.time() # Start timer

# Setting the parameters for model preprocessing and tuning from the caret package: 
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,           # 10-fold Crossvalidation
                           repeats = 5,           # repeated ten times
                           verboseIter=FALSE,
                           preProcOptions="pca",  # PCA Preprocessing
                           allowParallel=TRUE)    # With parallel backend

RF_model <- train(Class ~ ., 
                  data = training_set, 
                  method = "rf", 
                  preProcess = c("center","scale"), 
                  trControl= fitControl)

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```

#### Model output:
```{r res}
cache = TRUE
RF_model
```

#### View confusion matrix for model:
```{r cm}
cache = TRUE
RF_model$finalModel
```

#### Predict using the testing_set dataset
```{r predict}
cache = TRUE

#testing = as.factor(testing_set$Class)

#predictRF <- predict(modFit2, testing)

#confusionMatrix(predictRF, testing)
#confusionMatrixRF

#levels(predictRF)

#levels(testing_set)
```







Machine Settings:
```{r settings}
Sys.info()[c(1:3,5)]
```
```{r session}
sessionInfo()
```
EOF