---
title: "0-Convert-.FASTA-files-to-aa-strings.rmd
author: "Matthew Curcio"
date: "September 28, 2017"
output: html_document
---

## Title: 0-Convert-.FASTA-files-to-aa-strings.rmd

#### Summary: Converts "uniprot-([a-zA-Z]{3,4}).fasta" files to sequences of amino acids, one polypeptide per line. The files created are one file per protein classification.

| Protein Class  | Code Key |
|:---------------|:---------|
| HS-Non-O2      | Ctrl     |
| Erythrocruorin | Ery      |
| Hemerythrin    | Hhe      |
| Hemocyanin     | Hcy      |
| Leghemoglobin  | Lgb      |
| Myoglobin      | Mgb      |
| Hemoglobin     | Mgb      |

Libraries:
```{r libraries, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "seqinr", "stringr", "dplyr")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```

Read "uniprot-([a-zA-Z]{3,4}).fasta" files, convert and store sequences to '.txt'.
```{r}
read_fasta_2_txt <- function() {
    setwd("~/Dropbox/Oxy-RF/0_Data/1_FASTA-data")
    getwd()
    poly_peptides_files = list.files(pattern = ".fasta$")
    
    # loop to read sequences using seqinr:read.fasta
    for (i in seq(length(poly_peptides_files))) {
        prot_seq <- read.fasta(file = poly_peptides_files[i], 
                               seqtype = "AA",
                               seqonly = TRUE,
                               as.string = TRUE)
        file_name <- paste(poly_peptides_files[i], "_seq_ONLY.txt", sep = "")
        setwd("/home/mcc/Dropbox/git_projects/random_forest/1_cleaning_data")
        write.table(x = prot_seq, file = file_name, sep  = "\n",
                    row.names = FALSE, col.names = FALSE)
    }
}
```

