---
title: "Generate_500unit_test_harness.rmd"
author: "Matthew Curcio"
date: "March 4, 2019"
output: html_document
---

## Title: Generate_500unit_test_harness.rmd

#### Summary: 

It was seen through preliminary work that two problems became immediately appearent.  The first was that IF the files were concatenated they would something on the order of 26MB.  This 26MB file would be too large too run many manny trial of, therefore a test-harness was generated.  The second problem was the wide diversity of samples in each protein class.  The smallest grouping was Leghemoglobin, found in beans, which numbered only 125 proteins while the largest class was Hemerythrin with 40,763 proteins per class.

It was determined that the test-harness should have 500 units per class.  The 500 units would be chosen at random with Leghemoglobin being oversampled by 4:1. The largest class, Hemerythrin was undersampled, 81.25:1

This script will find the number of lines per file than sample 500 times from each file.  The lines will be .... (concat before or after?)

##### Libraries:
```{r libraries, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "seqinr", "stringr", "dplyr")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```



#### A. Create index of 500
```{r}
set.seed(1000)
generate_index <- function(line_count){
    index <- sample.int(line_count, n = 500, replace = TRUE)
}
```

#### B. 'Touch' file names for eventual 500 lines of Protein per class.
```{r}
write_empty_file_names <- function() {
    prot_list <- list.files(pattern = ".txt$")
    for (i in 1:length(prot_list)) {
        key <- gsub("^protein_class_([a-zA-Z]{3,4})\\.txt$", "\\1", prot_list[i])
        file_name <- paste(key, "_test_harness.lst", sep = "")
        fileConn <- file(file_name)
        writeLines(file_name, fileConn)
        close(fileConn)
    }
}
```

#### C. Find line count of 1 protein class file
```{r}
line_count <- length(readLines(prot_list[i]))
```


#### D. Transfer 1 line from protein class to Empty file class.
```{r}
#transfer_lines <- function(file_name) {
    con = file("protein_class_Ctrl.txt", "r")
    line = readLines(con, skip = 4, n = 1)
    print(line)
    #Function to wrtie to .lst file
    close(con)
#    }

```







Run fasta_2_sequences:
```{r}
start_time <- Sys.time() # Start timer
# Run function
end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```

Machine Settings:
```{r}
Sys.info()[c(1:3,5)]
```
```{r}
sessionInfo()
```
EOF