---
title: "050-ml-knn-caret-perc-aa-data.rmd"
author: "Matthew Curcio"
date: "February 20, 2019"
output: html_document
---

## Title: 

#### Summary: 


Libraries:
```{r libraries, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "ISLR", "caret", "doMC")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```

Import Data:
```{r, message=FALSE}
all_paa <- read_csv("/home/mcc/Desktop/ml-o2-binding-proteins/2-pipeline/Feb2019_protein_classes/all_paa.csv")
```

Simple Preparation:
```{r}
all_paa = all_paa[,-2]              # Remove TotalAA
#pro_class <- factor(all_paa$Class)
```


- Spliting data as training and test set. 
- Using createDataPartition() function from caret

createDataPartition:
```{r}
set.seed(1000)

indxTrain <- createDataPartition(y = all_paa$Class, 
                                 p = 0.75, 
                                 list = FALSE)
training <- all_paa[ indxTrain,]
testing  <- all_paa[-indxTrain,]

#Checking distibution in origanl data and partitioned data
prop.table(table(training$Class)) * 100
```


kNN requires variables to be normalized or scaled. caret provides software to preprocess data, therefore I am choosing centring & scaling

Preprocessing: "center", "scale"
```{r}
trainX <- training[ ,names(training)]
preProcValues <- preProcess(x = trainX, method = c("center", "scale"))
preProcValues
```

Training and train control:
```{r}
set.seed(1000)

start_time <- Sys.time() # Start timer
registerDoMC(cores = 3)

ctrl <- trainControl(method="repeatedcv",
                     repeats = 3)
knnFit <- train(Class ~ ., 
                data = training, 
                method = "knn", 
                trControl = ctrl, 
                preProcess = c("center","scale"), 
                tuneLength = 20)

end_time <- Sys.time()
end_time - start_time
```

Output of kNN fit:
```{r}
knnFit
```

# Plotting yields Number of Neighbours Vs accuracy (based on repeated cross validation)
```{r}
plot(knnFit, print.thres = 0.5, type = "S")
```


Get the confusion matrix to see accuracy value and other parameter values
```{r}
start_time <- Sys.time() # Start timer

knnPredict <- predict(knnFit, newdata = testing)
confusionMatrix(knnPredict, testing$Class)

end_time <- Sys.time()
end_time - start_time
```




























Machine Settings:
```{r}
Sys.info()[c(1:3,5,7)]
```

```{r}
sessionInfo()
```

EOF



















```{r}
sessioninfo()
```

```{r}
Uname????
```