---
title: "050-ml-knn-caret-perc-aa-data.rmd"
author: "Matthew Curcio"
date: "February 20, 2019"
output: html_document
---

http://rpubs.com/sediaz/tunelength_tunegrid
http://rpubs.com/jasonchanhku/loans
https://rpubs.com/ezgi/classification

http://rpubs.com/saqib/rpart  INTERESTING



## Title: C5.0 Trees

#### Summary: 

- Get the data
- Data Pre-processing
- Build a model
- Predictions
- Communication

Libraries:
```{r libraries, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "caret", "doMC", "ggplot2", "lattice", "C50")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```

Import Data:
```{r, message=FALSE}
X5_class_aa <- read_csv("2-pipeline/5_class_aa.csv")
```

Simple Preparation:
```{r}
five_paa = X5_class_aa[,-2]            # Remove TotalAA
class.factor <- factor(five_paa$Class) # 5 classes
```


- Spliting data as training and test set. 
- Using createDataPartition() function from caret

createDataPartition:
```{r}
set.seed(1000)

train_rows <- createDataPartition(y = five_paa$Class, 
                                  p = 0.75, 
                                  list = FALSE)
train_data <- five_paa[ train_rows,]
test_data  <- five_paa[-train_rows,]

#Checking distibution in origanl data and partitioned data
prop.table(table(train_data$Class)) * 100
```


Build a Decision Tree
Model the tree

Use Quinlan’s C5.0 decision tree algorithm implementation from the C50 package to build your decision tree
```{r}
c5_tree <- C5.0(train_data$Class ~ ., train_data)
```





















Machine Settings:
```{r}
Sys.info()[c(1:3,5,7)]
```

```{r}
sessionInfo()
```
EOF