---
title: "Heme AA Using KNN"
author: "Matthew Curcio"
date: "February 5, 2019"
output: html_document
---

From: https://rpubs.com/njvijay/16444


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Heme AA Using KNN

Import Heme Protein Data
```{r message=FALSE}
library(readr)
complete_aa <- read_csv("complete_aa.csv")
complete_aa <- complete_aa[,-2]
complete_aa <- factor(complete_aa$Class)
```

Caret is a valuable R package which was written in order to provide a general interface for nearly 150 ML algorithms. It also provides useful functions to produce training and testing datasets, preprocessing, evaluating the model, etc.,

```{r message=FALSE}
library(caret)
```

Split data into training and test sets. by using createDataPartition() function from caret
```{r}
set.seed(1000)

aa_train <- createDataPartition(y = complete_aa,
                                p = 0.8,
                                list = FALSE)
training <- complete_aa[ aa_train, ]
testing  <- complete_aa[-aa_train, ]
```


Check distibutions of original data and partitioned data
```{r}
prop.table(table(training$Class)) * 100
```

```{r}
prop.table(table(testing$Class)) * 100
```

This first trial will be carried without scaling or normalization since the values of AA composition are all <0.4.
```{r}
set.seed(2000)

start_time <- Sys.time()
ctrl <- trainControl(method = "repeatedcv",
                     repeats = 3)
end_time <- Sys.time()
end_time - start_time
```

K-nearest neighbor calculation
```{r}
start_time <- Sys.time()
knnFit <- train(Class ~ ., 
                data = training, 
                method = "knn", 
                trControl = ctrl, 
                tuneLength = 20)
end_time <- Sys.time()
end_time - start_time
```


```{r}
#Output of kNN fit
knnFit
```

```{r}
Libraries = c('lattice', 'ggplot2')
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```

```{r}
plot(knnFit)
```

Find confusion matrix to see accuracy value and other parameter values
```{r}
knnPredict <- predict(knnFit, testing)

confusionMatrix(knnPredict, testing$Class)
```



















