---
title: "031-EDA-part-ABC.rmd"
author: "Matthew Curcio"
date: "February 19, 2019"
output: html_document
---


## Title: 031-EDA-part-ABC.rmd

Summary: This script is the first of the exploratory data analysis for the oxygen binding proteins. Parts include:
- A. importing
- B. data types-str / inheritance / factors
  - I needed `tail -n +2 pro_class_Ctrl_aa.csv > ctrl_paa.csv` to delete first line of all .csv files, then use cat to splice them together leaving only ONE header notsix.
- C. head() / tail() commands
- D. Missing data exploration / Dealing w NA’s

#### Outline: 

```{r libraries}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c('readr')

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```

### A - Import Data Using "readr" package
```{r}
all_paa <- read_csv("2-pipeline/Feb2019_protein_classes/all_paa.csv")
```


### B - Data Types / inheritance / factors
```{r}
str(all_paa)
```


While importing the 28 MB dataset the 'Class' variable was properly imported as a character set. However we want to change this to a factor such that wwe can take advantage of R's classification abilities later on.

Make Column #1 factors
```{r}
factor_proteins = factor(all_paa$Class)
```

Now Check all levels
```{r}
levels(factor(all_paa$Class))  
```

```{r}
unique(all_paa$Class)
```


```{r}
table(all_paa$Class)
```

# PROBLEM??? with 6 Class names coming up????


### C. head() / tail() commands
```{r}
head (all_paa)
```

```{r}
tail (all_paa)
```

Let's look at the some specific lines
```{r}
all_paa[c(1215,1216,1217,1218,1219), ]
```


### D. Missing data exploration / Dealing w NA’s

Checking for missing data can be done in several ways. 

This gives us an indication that all columns are complete. In this case, command `is.na()` is a data type and in each column (e.g. 2) any are not `NA`.

We see there NO missing values NOW. So I will introduce some.

```{r}
apply(all_paa, 2, function(x) any(is.na(x)))
```

The `FALSE` indicates that there are no `NA`'s which appear in the dataset. 

--- 

#### Since our dataframe Does Not contain `NA` I will add some. 

When missing data is present. Generally there are several alternatives.

1. You can choose not use the column(feature) or row(observation) or that particular value in your computations.
2. Replace the missing values with the column average which is very commonly done.
3. Use a tool for data imputation, such as `mice`

I will demonsrate replacing the missing data with the column average.

```{r}
head(all_paa$TotalAA, 20)
```

Since our dataframe does not contain `NA` I will add some. 

I will arbitrarily introduce `NA` at values less than 500.

```{r}
all_paa$TotalAA[all_paa$TotalAA < 500] <- NA
head(all_paa$TotalAA, 20)
```



```{r}
head(is.na(all_paa$TotalAA), 20)
```

A useful function for finding `NA` values in the `any` and `all` functions.

```{r}
any(is.na(all_paa$TotalAA))  # Are ANY values `NA`
```

```{r}
all(is.na(all_paa$TotalAA))  # Are ALL values `NA`? Obviously not.
```

But where are the actual `NA` values in my column `all_paa$TotalAA`? 
```{r}
head(which(is.na(all_paa$TotalAA), arr.ind = TRUE))
```

The vector produced by `which` provides us with the position numbers for the `NA`s.

```{r}
mean_df = mean(all_paa$TotalAA, na.rm = TRUE)
mean_df
```

```{r}
NA2mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
all_paa[] <- lapply(all_paa, NA2mean)

head(all_paa$TotalAA, 20)
```

Do we see that positions (1, 2, 6, 8, 9, 10) are replaced by the mean?
mean


**END OF SECTION Explortory Data Analysis A-D



##### Machine Settings

```{r}
Sys.info()[c(1:3,5,7)]
```

```{r}
sessionInfo()
```

EOF



















```{r}
sessioninfo()
```

```{r}
Uname????
```