---
title: "032-EDA-part-EF.rmd"
author: "Matthew Curcio"
date: "February 19, 2019"
output: html_document
---

## Title: 032-EDA-part-EF.rmd

#### Summary: 

- E-Summary, Subsetting, Using *psych* package: `describeBy`
- F-1D Plots: Box plots / Violin plots

Libraries
```{r libraries, echo=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "psych", "ggplot2")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```

Import Data
```{r, message=FALSE}
all_paa <- read_csv("/home/mcc/Desktop/ml-o2-binding-proteins/2-pipeline/Feb2019_protein_classes/all_paa.csv")
Classes = factor(all_paa$Class)
```

### Descriptive Statistics

One can call any section or all of the features that are available.

For all: `summary(all_paa)` 

This command is overkill since it would produce 22 sets of information, one for every column.

**Remember:** It is possible to 'slice and dice' the data. For example, Data or dataframes (as they are refered to in R) are labeled:

    dataframe[row, columns]

Therefore below I wanted ony see the first 4 columns.
```{r}
summary(all_paa[, c(2,3,10,13)])
```

A very interesting library is the `psych` package. Not only does it have great datasets from [Sir Francis Galton](https://en.wikipedia.org/wiki/Francis_Galton), `cubits`,  `galton`, `heights` but some nice descriptive statistics as well, such as `describeBy`.

```{r}
describeBy(all_paa$TotalAA, group = Classes, mat=TRUE)
```


### 1D Plots Box

Box plots of Percent Amino Acid Composition

```{r fig.height=10, fig.width=6}
boxplot(all_paa[ ,c(3:22)],
        main = "Boxplot: Percent Composition of 20 AA",
        xlab = "Amino Acids",
        ylab = "Percent Composition",
        horizontal = TRUE,
        ylim = c(0,0.4))
```

1D Plots: Box plots of Polypeptide Length (# AA) Vs. Protein Class
```{r fig.height=4, fig.width=6}
boxplot(all_paa$TotalAA ~ Classes,
        main = "Boxplot: Number of AA Vs. Class",
        xlab = "Class",
        ylab = "Number of AA",
        horizontal = FALSE)
```

```{r fig.height=4, fig.width=6}
boxplot(all_paa$TotalAA ~ Classes,
        main = "Boxplot: Number of AA Vs. Class",
        xlab = "Class",
        ylab = "Number of AA",
        horizontal = FALSE,
        ylim = c(0, 1000))
```


1D Violin Plots
```{r}
ggplot(all_paa, aes(x=Classes, y=all_paa$TotalAA)) + geom_violin()
```


##### Machine Settings

```{r}
Sys.info()[c(1:3,5,7)]
```

```{r}
sessionInfo()
```

EOF