---
title: "033-EDA-part-GH.rmd"
author: "Matthew Curcio"
date: "February 19, 2019"
output: html_document
---

## Title: 033-EDA-part-GH.rmd

#### Summary: {-}

- G - Histograms
- H - 2D Plots: Scatter plots, Lattice / paired 2D plots
- I - Interactive Rotating 3D plots

Libraries:
```{r libraries, echo=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "lattice", "ggplot2")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```

Import Data:
```{r, message=FALSE}
all_paa <- read_csv("/home/mcc/Desktop/ml-o2-binding-proteins/2-pipeline/Feb2019_protein_classes/all_paa.csv")
Classes = factor(all_paa$Class)
```

### Histograms

This next secttion will explore the characteristics of the amino acid dataset with respect to 2D plots. This will allow us to look for correlations between features.

We will start with histograms, `hist()`:
```{r}
par(mfrow=c(1,2))
for (i in 3:4) {
    hist(all_paa[[i]],
         freq = FALSE,
         col = "pink",
         main = paste("% Probability Histogram of AA:" , colnames(all_paa)[i]),
         xlab = "%AA Composition",
         ylab = "% Probability",
         xlim = c(0, 0.4),
         ylim = c(0,20))
}
```

---

#### Simple Plot Options {-}

| Description | Example |
|-------------|---------|
| Main Title & Sub-Titles |`main = "TITLE", sub = "subtitle"` |
| X & Y<br>Axis Labels |`xlab = "bottom", ylab = "left-side"` |
| X & Y<br>Axis Limits |`xlim = c(0,10), ylim = c(0,10)`<br><br>If no limits are set, default values will be used.<br>. |
| Use Log scales |`xlog = TRUE, ylog = TRUE` |
| Line Type |  [img[lty = 1:6|http://www.statmethods.net/advgraphs/images/lines.png]]<br>OR TRY:<br>`plot (child, parent, type="l", lty=0:6)` |
| Line Width |`lwd = 2` |
| Color |`col = ff0000` |
| ''pch'' Magnification Factor  |`cex = 2` (default = 1) |

---


Simple Scatterplot
```{r}
plot(x = all_paa[[3]], 
     y = all_paa[[4]], 
     main = "Simple Scatterplot of G vs. P",
     xlab = "AA: G", 
     ylab = "AA: P", 
     pch = 4) 
```

## Move to Appendix
## If you are looking for different symbols to use if is easy enough to add a simple chart.

```{r}
plot(0:25, pch=0:25)
```

### Quantile-Quantile plot 

QQ plots are useful for determining normality:
```{r}
par(mfrow=c(1,2))
for (i in 3:4) {
    qqnorm(all_paa[[i]],
           main = paste("QQ Plot of AA:" , colnames(all_paa)[i]))
    qqline(all_paa[[i]])
}
```

By the looks at these graphics, the distributions of amino acids is NOT normal. To determine the correctt distrubution would take more investigation.

---























##### Machine Settings

```{r}
Sys.info()[c(1:3,5,7)]
```

```{r}
sessionInfo()
```

EOF