---
title: "O2-Binders-Missing-Data"
author: "Matthew Curcio"
date: "February 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## O2-Binders-Missing-Data

This section will explore situations where data is missing from a dataframe.

There are several very simple options that will fill in these missing values.

1. Importing Data Using 'readr' package.

```{r message=FALSE, include=FALSE}
setwd("~/Desktop/ml-o2-binding-proteins/2-pipeline/7-percent-aa-composition-numbers")
library(readr)
df <- read_csv("complete_aa.csv")
```


2. Checking for missing data can be done in several ways. 

This gives us an indication that all columns are complete. In this case, command `is.na()` is a data type and in each column (e.g. 2) any are not `NA`.

We see there NO missing values NOW. So I will introduce some.

```{r}
apply(df, 2, function(x) any(is.na(x)))
```

Since our dataframe Does Not contain `NA` I will add some. 

When missing data is present. Generally there are several alternatives.

1. You can choose not use the column(feature) or row(observation) or that particular value in your computations.
2. Replace the missing values with the column average. This is very commonly done.
3. Use a tool for data imputation.

I will demonsrate replacing the missing data with the column average.

```{r}
head(df$TotalAA, 20)
```

Since our dataframe does not contain `NA` I will add some. 

I will arbitrarily introduce `NA` at values less than 500.

```{r}
df$TotalAA[df$TotalAA < 500] <- NA
head(df$TotalAA, 20)
```



```{r}
head(is.na(df$TotalAA), 20)
```

A useful function for finding `NA` values in the `any` and `all` functions.

```{r}
any(is.na(df$TotalAA))  # Are ANY values `NA`
```

```{r}
all(is.na(df$TotalAA))  # Are ALL values `NA`? Obviously not.
```

But where are the actual `NA` values in my column `df$TotalAA`? 
```{r}
head(which(is.na(df$TotalAA), arr.ind = TRUE))
```

The vector produced by `which` provides us with the position numbers for the `NA`s.

```{r}
mean_df = mean(df$TotalAA, na.rm = TRUE)
mean_df
```

```{r}
NA2mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
df[] <- lapply(df, NA2mean)

head(df$TotalAA, 20)
```

Do we see that positions (1, 14, 15, 19) are replaced by the mean?
mean
















