---
title: "O2-Binders-EDA"
author: "Matthew Curcio"
date: "February 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# O2-Binders-EDA


cat *.csv > complete_aa.csv
```{r include=FALSE}
setwd("~/Desktop/ml-o2-binding-proteins/2-pipeline/7-percent-aa-composition-numbers")
library(readr)
protein_complete_aa <- read_csv("complete_aa.csv")
```

```{r}
p_aa <- protein_complete_aa[-2]
factor_aa = factor(p_aa$Class)
head(p_aa)
```

```{r}
tail(p_aa)
```

Check for any missing data
```{r}
apply(p_aa, 2, function(x) any(is.na(x)))
```

Produce table of protein types
```{r}
table(p_aa$Class)
```

```{r}
prop.table(table(p_aa$Class))
```

Summary Statistics

```{r}
sum_p_aa = p_aa[-1] # Delete the 'Class' feature since that is a factor value(0:6)
summary(sum_p_aa)
```

```{r}
fivenum(sum_p_aa$G)
```


OK, Let's keep going with EDA

```{r fig.height=11}
boxplot(sum_p_aa, 
        horizontal = TRUE, 
        main = "Percent AA Composition Oxygen-Binding Proteins")
```

## DataExplorer

Using the DataExplorer package we can a different view.

```{r}
library(DataExplorer)

plot_missing(p_aa)
```


```{r fig.height=8, fig.width=8}
plot_correlation(p_aa, 
                 type="all", 
                 title = "Correlation Plot of Percent AA Compostion")
```

## Compute the correlation matrix for the 20 AA variables

```{r message=FALSE}
library(corrplot) # produces correlation plots of variables
```

```{r}
# Reduce total_aa data set to amino acids only NOT Class
corrSet <- p_aa[,2:21]

str(corrSet)
```


```{r}
corrMatrix <- cor(corrSet) # Calculate correlation coefficients
#corrMatrix
```

Generate (Upper) Correlation plot of Amino Acids

- Combine correlogram with the significance test
- Mark the insignificant coefficients according to the specified p-value significance level with ‘x’.

```{r message=FALSE}
library(Hmisc)
```

```{r}
cor_aa <- rcorr(as.matrix(corrMatrix))
M <- cor_aa$r
p_mat <- cor_aa$P

corrplot(corrMatrix, 
         method = "color", 
         type = "upper", 
         order = "hclust", 
         p.mat = p_mat, 
         sig.level = 0.01)
text(9, 1.7, "Correlation Plot (R^2) of AA", cex = 2)
text(9, 4, "'X' indicate insignificant p-value (p<0.01) for Correlation")
```

## Display the Sigificant Correlation Coefficients Only:
While leaving the insigificant values blank

```{r}
corrplot(corrMatrix, 
         method = "number", 
         type = "upper", 
         order = "hclust",
         p.mat = p_mat, 
         sig.level = 0.01, 
         insig = "blank")
text(9, 1.7, "Correlation Value Plot (R^2) of AA", cex = 2)
text(9, 3, "Numbered squares indicate significant Correlation Values given p=0.01")
```

The results from the Correlation Matrix provides us with 10 amino acids that seem to be correlated.

These are: 

G, P, A, I, N, K, E, M, F, Y
OR
Gly, Pro, Ala, Ile, Asn, Lys, Met, Phe, Tyr


---

# TO DO:
1. Boxplots of Total AA Versus Protein Category
2. Boxplots of M.W. of Protein Categories
3. Violin plots 
4. ?

https://rpubs.com/oaxacamatt/exploratory_MONO


Training models:
https://rpubs.com/oaxacamatt/CART_20AA



### Machine Set up
```{r}
sessionInfo()
```

```{r}
Sys.info()[c(1:3,5,7)]
```

EOF