---
title: "EDA-Part1-Oxygen-binding.rmd"
author: "Matthew Curcio"
date: "March 18, 2019"
output: html_document
---

## File Name: EDA-Part1-Oxygen-binding.rmd

### Summary: 
One important question which I would like to answer with an exploratory
data analysis is this data set sufficient to succesfully classify the seven classes of protein using a Random Forest machine learning approach? This should clearer as we test and further describe the data.

| Code Key | Protein Class  |
|:---------|:---------------|
| 0 / Ctrl | Controls consist of human proteins which do not bind oxygen |
| 1 / Ery  | Erythrocruorin |
| 2 / Hcy  | Hemocyanin     |
| 3 / Hgb  | Hemoglobin     |
| 4 / Hhe  | Hemerythrin    |
| 5 / Lgb  | Leghemoglobin  |
| 6 / Mgb  | Myoglobin      |


Libraries:
```{r libraries, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "psych", "knitr")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```


#### A. Load Data
Import Single Amino Acid percent composition for first round of Exploratory Data Analysis.
```{r load_data, message=FALSE, warning=FALSE, include=TRUE}
complete_aa <- read_csv("complete_aa.csv")
```


#### B. Check Data Head
```{r}
head(complete_aa)
```


#### C. Check Data Tail
```{r}
tail(complete_aa)
```


#### D. Find Dimensions:
The dataset consists of 22 variables and 16,952 observations, consisting of 7 classes, (0:6).
```{r dim}
dim(complete_aa)
```


#### E. Examine Data Structure: 
Note: Class is a numerical value and must be converted to a categorical before analysis.
```{r str}
str(complete_aa)
```


#### F. Convert Class To Factor
Convert Class(numerical) to Factor of 7 Protein Classes(Prot_Class)
```{r factor}
prot_class <- as.factor(complete_aa$Class)

typeof(complete_aa$Class)
class(complete_aa$Class)
```


#### G. Produce Summary
Produce Summary information for all 22 columns.
The `summary` command produces a 6 number summary.
```{r summary}
summary(complete_aa$TotalAA)
```


#### H. Using Package "psych::describe" commands
Using 'describe' we are able to investigate the first four moments. 

For brevity, we will calculate 2 amino acids (R, H)

| Single Letter | Amino Acid |
|:--------------|:-----------|
| R             | Arginine   |
| H             | Histidine  |


```{r arginine}
describe(complete_aa$R)
```

```{r histidine}
describe(complete_aa$H)
```


#### I. Testing for Missing Data

In R, missing values are represented by 'NA' (not available). 
This is not to be confused with impossible values (e.g., dividing by zero) are represented by 'NaN' (not a number). 
```{r}
kable(sapply(complete_aa, function(x) sum(is.na(x))))
```
There are NO missing values in any of the 22 features.


#### J. Principle Component Analysis: `princomp`

`princomp` has a component called `sdev` that is the "Standard deviation" for the dataframes features, therefore we can calculate variance.
```{r}
aa_PCA = princomp(complete_aa[,3:22])

aa_PCA$sdev^2
```

The proportion of variance is the variance divided by the sum of all variances. 
```{r include=FALSE}
aa_PCA$sdev^2 / sum(aa_PCA$sdev^2)
```

The Cumulative Proportion is the cumulative sum of the proportion of variance.
```{r}
cumsum(aa_PCA$sdev^2 / sum(aa_PCA$sdev^2))
```

Now you have the Cumulative Proportion values, just plot them. 
```{r}
plot(cumsum(aa_PCA$sdev^2 / sum(aa_PCA$sdev^2)), 
     main = "Cumulative Proportion of varainces of all 20 Amino Acids(AA)",
     xlab = "Principal Component",
     ylab = "Proportion of AA",
     ylim = c(0,1),
     type="b")
```

The contribution of each amino acid is hard to discern from this graphic alone, therefore  I will use another library to investigate further.

---

#### K. Machine Settings:
```{r}
Sys.info()[c(1:3,5)]
```
```{r}
sessionInfo()
```
EOF
