---
title: "EDA-Part2-Oxygen-binding.rmd"
author: "Matthew Curcio"
date: "March 18, 2019"
output: html_document
---

## File Name: EDA-Part2-Oxygen-binding.rmd

### Summary: 
One important question which I would like to answer with an exploratory
data analysis is this data set sufficient to succesfully classify the seven classes of protein using a Random Forest machine learning approach? This should clearer as we test and further describe the data.

| Code Key | Protein Class  |
|:---------|:---------------|
| 0 / Ctrl | Controls consist of human proteins which do not bind oxygen |
| 1 / Ery  | Erythrocruorin |
| 2 / Hcy  | Hemocyanin     |
| 3 / Hgb  | Hemoglobin     |
| 4 / Hhe  | Hemerythrin    |
| 5 / Lgb  | Leghemoglobin  |
| 6 / Mgb  | Myoglobin      |


Libraries:
```{r libraries, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "knitr")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```


#### A. Load Data
Import Single Amino Acid percent composition for first round of Exploratory Data Analysis.
```{r load_data, message=FALSE, warning=FALSE, include=TRUE}
complete_aa <- read_csv("complete_aa.csv")
```


#### B. Number of Proteins Per Class

```{r}
class_table <- table(complete_aa$Class)
knitr::kable(class_table)
```


#### C. Plot Variances vs Amino Acid

```{r}
aa_vars = apply(complete_aa[,3:22], 2, var)
aa_vars_sorted = sort(10^5*aa_vars, decreasing = TRUE)
aa_vars_sorted
```


#### D. Plot of Variances
```{r}
aa_names <- c("P", "L", "K", "E", "G", 
              "A", "S", "R", "C", "Q", 
              "I", "V", "T", "D", "F", 
              "N", "H", "Y", "M", "W") 

plot(aa_vars_sorted,
     main = "Plot of Variances (x 10^5) Vs Amino Acid Type",
     ylab = "Variances (x 10^5)",
     xlab = "Amino Acid Type",
     ylim =  c(0,100),
     xaxt = "n")
axis(1, at = 1:20, labels = aa_names)
```


#### E. % Means vs Amino Acids
```{r}
aa_means = apply(complete_aa[,3:22], 2, mean)
aa_means_sorted = sort(100*aa_means, decreasing = TRUE)
aa_means_sorted
```


#### F. Plot of Means (% of Total) Vs Amino Acid Type 
```{r}
plot(aa_means_sorted,
     main = "Plot of Means (% of Total) Vs Amino Acid Type",
     ylab = "Means % of Total)",
     xlab = "Amino Acid Type",
     ylim =  c(0,10),
     xaxt = "n")
axis(1, at = 1:20, labels = aa_names)
```

---

#### G. Machine Settings:
```{r}
Sys.info()[c(1:3,5)]
```
```{r}
sessionInfo()
```
EOF
