---
title: "Exploratory Data Analysis on MONO Peptides"
author: "Matthew Curcio"
date: "October 3, 2017"
output: html_document
---

```{r library}
library(readr)
```

protein_class = c("homosapiens", "erythrocruorin", "hemerythrin",
                      "hemocyanin", "leghemoglobin", "myoglobin",
                      "hemoglobin")

```{r homosapiens}
protein_class_0_aa <- read_csv("~/Dropbox/git_projects/random_forest/4_exploratory/mono/protein_class_0_aa.csv")

summary(protein_class_0_aa)
```

```{r erythrocruorin}
protein_class_1_aa <- read_csv("~/Dropbox/git_projects/random_forest/4_exploratory/mono/protein_class_1_aa.csv")

summary(protein_class_1_aa)
```


```{r}
protein_class_2_aa <- read_csv("~/Dropbox/git_projects/random_forest/4_exploratory/mono/protein_class_2_aa.csv")

summary(protein_class_2_aa)
```



```{r}
protein_class_3_aa <- read_csv("~/Dropbox/git_projects/random_forest/4_exploratory/mono/protein_class_3_aa.csv")

summary(protein_class_3_aa)
```



```{r}
protein_class_4_aa <- read_csv("~/Dropbox/git_projects/random_forest/4_exploratory/mono/protein_class_4_aa.csv")

summary(protein_class_4_aa)
```



```{r}
protein_class_5_aa <- read_csv("~/Dropbox/git_projects/random_forest/4_exploratory/mono/protein_class_5_aa.csv")

summary(protein_class_5_aa)
```



```{r}
protein_class_6_aa <- read_csv("~/Dropbox/git_projects/random_forest/4_exploratory/mono/protein_class_6_aa.csv")

summary(protein_class_6_aa)
```

## Boxplots of Total AA Versus Protein Category

```{r}
setwd("~/Dropbox/git_projects/random_forest/data")

boxplot(protein_class_0_aa$TotalAA, protein_class_1_aa$TotalAA,
        protein_class_2_aa$TotalAA, protein_class_3_aa$TotalAA,
        protein_class_4_aa$TotalAA, protein_class_5_aa$TotalAA,
        protein_class_6_aa$TotalAA,
        log = "y", horizontal = FALSE,
        names = c("homosapiens", "erythrocruorin", "hemerythrin",
                  "hemocyanin", "leghemoglobin", "myoglobin", "hemoglobin"),
        main = "Total AA Versus Protein Category",
        ylab = "Total Amino Acids per Protein")

# Save copy to file.
svg(filename="boxplot_total_aa_vs_protein_category.svg")
boxplot(protein_class_0_aa$TotalAA, protein_class_1_aa$TotalAA,
        protein_class_2_aa$TotalAA, protein_class_3_aa$TotalAA,
        protein_class_4_aa$TotalAA, protein_class_5_aa$TotalAA,
        protein_class_6_aa$TotalAA,
        log = "y", horizontal = FALSE,
        names = c("homosapiens", "erythrocruorin", "hemerythrin",
                  "hemocyanin", "leghemoglobin", "myoglobin", "hemoglobin"),
        main = "Total AA Versus Protein Category",
        ylab = "Total Amino Acids per Protein")
dev.off()

```

### Due to the very low molecular weight proteins drastically changing the percent amino acid composition of some of the categories of proteins I decided to trim proteins below 26 AA in length from three categories;

### After looking at the Hemoglobin file morre closely I found a protein with a count of 1156 which is not part of the normal grouping of hemoglobins.  Therefore I have trimmed this single protein as well.

* homosapiens, class = 0
* hemocyanin, class = 3
* hemoglobin, class = 6

```{r trimming_datasets}
setwd("~/Dropbox/git_projects/random_forest/2_single_aa")

trimmed_homo_sapien_class_0 = subset(protein_class_0_aa, protein_class_0_aa[,2] > 27)
write.csv(x = trimmed_homo_sapien_class_0, 
          file = "trimmed_homo_sapien_class_0.csv",
          row.names = FALSE)

trimmed_hemocyanin_class_4 = subset(protein_class_4_aa, protein_class_4_aa[,2] > 27)
write.csv(x = trimmed_hemocyanin_class_4, 
          file = "trimmed_hemocyanin_class_4.csv",
          row.names = FALSE)

trimmed_hemoglobin_class_6 = subset(protein_class_6_aa, protein_class_6_aa[,2] > 27 &
                                    protein_class_6_aa[,2] < 1155)
write.csv(x = trimmed_hemoglobin_class_6, 
          file = "trimmed_hemoglobin_class_6.csv",
          row.names = FALSE)
```


```{r}
setwd("~/Dropbox/git_projects/random_forest/data")

boxplot(trimmed_homo_sapien_class_0$TotalAA, protein_class_1_aa$TotalAA,
        protein_class_2_aa$TotalAA, protein_class_3_aa$TotalAA,
        trimmed_hemocyanin_class_4$TotalAA, protein_class_5_aa$TotalAA,
        trimmed_hemoglobin_class_6$TotalAA,
        log = "y", horizontal = FALSE,
        names = c("homosapiens", "erythrocruorin", "hemerythrin",
                  "hemocyanin", "leghemoglobin", "myoglobin", "hemoglobin"),
        main = "Trimmed Total AA Versus Protein Category",
        ylab = "Total Amino Acids per Protein")

# Save copy to file.
svg(filename="boxplot_total_aa_vs_protein_category.svg")
boxplot(trimmed_homo_sapien_class_0$TotalAA, protein_class_1_aa$TotalAA,
        protein_class_2_aa$TotalAA, protein_class_3_aa$TotalAA,
        trimmed_hemocyanin_class_4$TotalAA, protein_class_5_aa$TotalAA,
        trimmed_hemoglobin_class_6$TotalAA,
        log = "y", horizontal = FALSE,
        names = c("homosapiens", "erythrocruorin", "hemerythrin",
                  "hemocyanin", "leghemoglobin", "myoglobin", "hemoglobin"),
        main = "Trimmed Total AA Versus Protein Category",
        ylab = "Total Amino Acids per Protein")
dev.off()
```

```{r}
summary(trimmed_homo_sapien_class_0)
```


```{r}
summary(trimmed_hemocyanin_class_4)
```


```{r}
summary(trimmed_hemoglobin_class_6)
```

```{r concatenate_mono_files}

all_proteins_mono <- rbind.data.frame(trimmed_homo_sapien_class_0, 
                                      protein_class_1_aa,
                                      protein_class_2_aa, 
                                      protein_class_3_aa,
                                      trimmed_hemocyanin_class_4,
                                      protein_class_5_aa,
                                      trimmed_hemoglobin_class_6)

# Change Protein Categories to ether (0,1) depending on whether they are oxygen binding or not.

for (i in length(all_proteins_mono)) {
    if (all_proteins_mono[i,1] > 1) {
        # Change all categories to either 0 or 1
        all_proteins_mono[i,1] = 1
    }
}

setwd("~/Dropbox/git_projects/random_forest/5_randomForest")
write.csv(x = all_proteins_mono, 
          file = "trimmed_all_proteins_mono.csv",
          row.names = FALSE)
```

