---
title: "Generate_500unit_test_harness.rmd"
author: "Matthew Curcio"
date: "March 4, 2019"
output: html_document
---

## Title: Generate_500unit_test_harness.rmd

#### Summary: 

It was seen through preliminary work that two problems became immediately appearent.  The first was that IF the files were concatenated they would something on the order of 26MB.  This 26MB file would be too large too run many manny trial of, therefore a test-harness was generated.  The second problem was the wide diversity of samples in each protein class.  The smallest grouping was Leghemoglobin, found in beans, which numbered only 125 proteins while the largest class was Hemerythrin with 40,763 proteins per class.

It was determined that the test-harness should have 500 units per class.  The 500 units would be chosen at random with Leghemoglobin being oversampled by 4:1. The largest class, Hemerythrin was undersampled, ~1:81.25

This script will find the number of lines per file than sample 500 times from each file.  The lines will be .... (concat before or after?)

Libraries:
```{r libraries, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Libraries = c("readr", "seqinr", "stringr", "dplyr")

# Install if not present
for(p in Libraries){
    if(!require(p, character.only = TRUE))
        install.packages(p)
    library(p, character.only = TRUE)
}
```


Read from the index then Readline from .txt file and write to test-harness.
```{r}
generate_index <- function(input_file){
    line_count <- length(readLines(input-file))
    print(paste(input_file, "=", line_count, "lines.", sep = " "))
    index <- sample.int(500, n = line_count, replace = TRUE)
    return(index)
}

generate_test_files <- function(file_input){
    
}

loop_thru_files <- function(){
    setwd()
    prot_list <- list.files(pattern = ".txt$")
    #print(paste("Input", length(prot_list), "files:", sep=" "))
    #print(prot_list)
    for (i in 1:length(prot_list)) {
        #print(paste("Processing: ", prot_list[1], sep = ""))
        
        generate-test-files(prot_list[i])
}
```


```{r}
#timestamp_header <- function(){
    
#}
run_time <- as.POSIXlt(Sys.time())
timestamp_header <- paste("A","-",
                          run_time$year,"-",
                          run_time$mon,"-",
                          run_time$mday,"-",
                          "T",
                          run_time$hour,"-",
                          run_time$min,
                          sep = "")

timestamp_header
```

















```{r}
processFile = function(filename) {
    on = file(filename, "r")
    while ( TRUE ) {
        line = readLines(con, n = 1)
        if ( length(line) == 0 ) {
            break
            }
        print(line)
        }
    close(con)
    }

choose_lines_from_index() <- function(index){
    for (j in 1:500) {
        index[j] 
        
    }
}
```


```{r}
set.seed(1000)

find_line_counts <- function(){
    #AA_DIR <- "/OxyRF/0_Data/aa-strings-data"
    #setwd(AA_DIR)
    #setwd("~/Dropbox/Oxy-RF/0_Data/aa-strings-data")
    
        
        key_code <- sub("protein_class_(.*)\\.txt", "\\1", prot_list[i])
        file_type <- paste(key_code, "_test_harness.txt", sep = "")
        
        file_connect<-file(type_timestamp_line)
        writeLines(type_timestamp_line, file_connect)
        
        #choose_lines_from_index(index)
        #return(index)
    }
}
```


```{r}
find_line_counts()
```






Run fasta_2_sequences:
```{r}
start_time <- Sys.time() # Start timer

# Run function


end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```



Machine Settings:
```{r}
Sys.info()[c(1:3,5)]
```
```{r}
sessionInfo()
```
EOF